```{r micro,results = "hide"}
source("../R/01_00_load_data.R")
```

## Bootstrap estimate of SE

Use the bootstrap to estimate the standard error of $\hat\theta = \mu_A - \mu_B$, where $\mu_A$ and $\mu_B$ are sample means of counts for methods A and B, respectively. Use $B = 2000$ bootstrap samples. Assume the replicates are **unpaired**.  

  * Let $A=X$ and $Y=B$. I am changing the labeling so as not to confuse with the number of bootstrap samples, $B$.
  * sampling is unpaired
  * Poisson distribution is assumed. Hence, this is parametric bootstrap. Recall:
  
  \begin{equation}
  \begin{aligned}
  \mathcal P(X = x \mid \lambda) = \frac{e^{-\lambda} \lambda^x}{x!}; \quad x = 0, 1, \dots; \quad 0 \leq \lambda < \infty
  \end{aligned}
  \end{equation}

  * Following a fact mentioned on page 237 from @efron, a sufficient statistic for the parameter of the Poisson distribution $\lambda$ is the average of the observations $\sum \frac{x}{n}$.

__Algorithm:__

**Step 1**: Calculate a sufficient statistic for $\lambda_X$ (method A) and $\lambda_Y$ (method B).

\begin{equation}
\begin{aligned}
\hat \lambda_X = \frac {\sum_{i = 1}^n x_i}{n} \\
\hat \lambda_Y = \frac {\sum_{i = 1}^n y_i}{n}
\end{aligned}
\end{equation}  


**Step 2**: Let $B$ be the number of bootstrap samples taken. With $n =$ `r n` and for each method, sample from the plug-in distribution.


\begin{equation}
\begin{aligned}
B^*_{Y,1} &= \left\{y_{1,1}^*, \dots, y_{1,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_Y\right)\\
B^*_{Y,2} &= \left\{y_{2,1}^*, \dots, y_{2,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_Y\right)\\
\vdots &= \vdots \\
B^*_{Y,2000} &= \left\{y_{2000,1}^*, \dots, y_{2000,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_Y\right)\\
\end{aligned}
\end{equation}  


\begin{equation}
\begin{aligned}
B^*_{X,1} &= \left\{x_{1,1}^*, \dots, x_{1,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_X\right)\\
B^*_{X,2} &= \left\{x_{2,1}^*, \dots, x_{2,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_X\right)\\
\vdots &= \vdots \\
B^*_{X,2000} &= \left\{x_{2000,1}^*, \dots, x_{2000,8}^*\right\} \overset{iid}\sim \mathcal P\left(\hat \lambda_X\right)\\
\end{aligned}
\end{equation}  


**Step 3**: For each of the bootstrap samples, calculate $\hat\theta^*_b = \hat \mu_{X,b}^* - \hat \mu_{Y,b}^*; \quad b = 1, \dots 2000$.

**Step 4**: Calculate

\begin{equation}
\begin{aligned}
\widehat {se} \left(\hat \theta\right) = \sqrt \frac{\sum_{b=1}^B \left(\hat \theta_b^*-\bar{\hat \theta}^* \right)^2}{B-1};\ \ \bar{\hat \theta}^* = \frac{1}{B}\sum^B_{b=1} \hat \theta^*_b
\end{aligned}
\end{equation}  


```{r se}
R01_01 <- "../R/01_01_se_estimate.R"
source(R01_01)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_01, paste(rep("-", 41), collapse = "")))
cat(paste(readLines(R01_01), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

The bootstrap estimate of the standard error is `r suff_se`.


## Histogram

The histogram has a symmetric bell curve shape. This is expected to be centered on the statistic. In this case the statistic $\hat \theta$ is `r mean(microbiology$X)-mean(microbiology$Y)` while the mean of $\hat \theta^*$'s is `r mean(suff_theta_star)`.

```{r echo = FALSE}
hist(
  suff_theta_star,
  main = expression(paste("Histogram of ", hat(theta), "*")),
  xlab = expression(paste(hat(theta), "*")),
  )
```

## Confidence Intervals

We are $95\%$ confident that the true value of $\theta$ is between `r suff_ci[1]` and `r suff_ci[2]`. From my perspective, the interval is broad since the difference ranges from almost zero to a relatively large positive number. However, this gives us an idea that the difference is positive.


