---
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    toc: false
  citation_package: natbib
fontsize: 12pt 
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \usepackage{setspace}\singlespacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{indentfirst}
- \usepackage{titling}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[RO,RE]{Matala}
- \fancyhead[LO,LE]{Midterm}
- \fancyfoot[LO,LE]{`r format(Sys.time(), 'Last rendered on %Y %B %d %X')`}
- \fancyfoot[CO,CE]{}
- \fancyfoot[RE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
geometry: margin=2cm
papersize: a4
bibliography: /home/scientists/sci01/Projects/bootstrap/common/ref.bib
biblio-style: apalike
csl: /home/scientists/sci01/Projects/bootstrap/common/harvard-educational-review.csl
nocite: |
  @efron
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4,
  fig.width = 5)
options(scipen = 1, digits = 7)
```

# Microbiology

A laboratory wants **to determine if two different methods (A and B) give similar results** for quantifying a particular bacterial species in a particular medium. Under each method, the counts form a random sample. **Assume that the counts follow a Poisson distribution**, since this distribution is a typical model for such data. Let
$\mu_A$ and $\mu_B$ represent the population mean counts for Methods A and B, respectively. Let $\theta = \mu_A -\mu_B$.


```{r, child = "child/item_01.Rmd"}

```

# Fishery

The dataset fishery.csv contains 40 annual counts of the numbers of recruits (R) and spawners (S) in a salmon population. The **units are in thousands of fish**. Recruits are fish that enter the catchable population. Spawners are fish that are laying eggs. Spawners die after laying eggs. The classic model for the relationship between spawners and recruits is

$$
R = \frac{1}{\beta_0  + \frac{\beta_1}{S}}; \quad \beta_0 \geq 0 \text{ and }\beta_1 \geq 1
$$

We can fit such a model by using a linear regression given by

$$
\frac{1}{R_i} = \beta_0 + \beta_1 \frac{1}{S_i} + \epsilon_i; \quad i = 1, \dots, 40
$$

The **S variable can be considered as fixed**. Suppose that fish are iid with mean 0 and finite variance, but their **distribution is unknown**. The total population abundance will only stabilize if $R = S$. Thus, the stable population level is the point where the line $R = S$ intersects the curve relating R and S. The total population will decline if fewer recruits are produced than the number of spawners who died producing them. If too many recruits are produced, the population will also decline eventually because there is not enough food for them all. Thus, only some middle level of recruits can be sustained indefinitely in a stable population. Let $S_0$ be the value of $S$ at the stable population level.

```{r, child = "child/item_02.Rmd"}

```

# References

<div id="refs"></div>

# Appendix

```{r, child = "child/appendix.Rmd"}

```