---
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    toc: false
  citation_package: natbib
fontsize: 12pt 
indent: true
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \usepackage{setspace}\singlespacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{indentfirst}
- \usepackage{titling}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[RO,RE]{Matala}
- \fancyhead[LO,LE]{Midterm}
- \fancyfoot[LO,LE]{`r format(Sys.time(), 'Last rendered on %Y %B %d %X')`}
- \fancyfoot[CO,CE]{}
- \fancyfoot[RE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
geometry: margin=2cm
papersize: a4
bibliography: /home/scientists/sci01/Projects/bootstrap/common/ref.bib
biblio-style: apalike
csl: /home/scientists/sci01/Projects/bootstrap/common/harvard-educational-review.csl
nocite: |
  @efron
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4,
  fig.width = 5)
options(scipen = 1, digits = 5)
```

# Microbiology

A laboratory wants **to determine if two different methods (A and B) give similar results** for quantifying a particular bacterial species in a particular medium. Under each method, the counts form a random sample. **Assume that the counts follow a Poisson distribution**, since this distribution is a typical model for such data. Let
$\mu_A$ and $\mu_B$ represent the population mean counts for Methods A and B, respectively. Let $\theta = \mu_A -\mu_B$.


```{r, child = "child/item_01.Rmd"}

```

# Fishery

The dataset fishery.csv contains 40 annual counts of the numbers of recruits (R) and spawners (S) in a salmon population. The units are in thousands of fish. Recruits are fish that enter the catchable population. Spawners are fish that are laying eggs. Spawners die after laying eggs. The classic model for the relationship between spawners and recruits is

$$
R = \frac{1}{\beta_0  + \frac{\beta_1}{S}}; \quad \beta_0 \geq 0 \text{ and }\beta_1 \geq 1
$$

We can fit such a model by using a linear regression given by

$$
\frac{1}{R_i}
$$

# References

<div id="refs"></div>

# Appendix

```{r, child = "child/appendix.Rmd"}

```