---
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    toc: false
  citation_package: natbib
fontsize: 12pt 
indent: true
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \usepackage{setspace}\singlespacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{indentfirst}
- \usepackage{titling}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[RO,RE]{Matala}
- \fancyhead[LO,LE]{Midterm}
- \fancyfoot[LO,LE]{`r format(Sys.time(), 'Last rendered on %Y %B %d %X')`}
- \fancyfoot[CO,CE]{}
- \fancyfoot[RE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
geometry: margin=2cm
papersize: a4
bibliography: /home/scientists/sci01/Projects/bootstrap/common/ref.bib
biblio-style: apalike
csl: /home/scientists/sci01/Projects/bootstrap/common/harvard-educational-review.csl
nocite: |
  @sonnet
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4,
  fig.width = 5)
options(scipen = 1, digits = 5)
```

# Microbiology

A laboratory wants **to determine if two different methods(A and B) give similar results** for quantifying a particular bacterial species in a particular medium. Under each method, the counts form a random sample. **Assume that the counts follow a Poisson distribution**, since this distribution is a typical model for such data. Let
$\mu_A$ and $\mu_B$ represent the population mean counts for Methods A and B, respectively. Let $\theta = \mu_A \mu_B$.


```{r, child = "child/item_01.Rmd"}

```

# Fishery

# References

<div id="refs"></div>

# Appendix

```{r, child = "child/appendix.Rmd"}

```