## Algorithm

**Step 1**: Let $B$ be the number of bootstrap samples taken. With $n = 15$, do SRSWR from schools 1 to 15.


\begin{equation}
\begin{aligned}
B^*_1 &= \left\{(X^*_{11}, Y^*_{11}), (X^*_{21}, Y^*_{21}), \ldots, (X^*_{n1}, Y^*_{n1})\right\} \\
B^*_2 &= \left\{(X^*_{12}, Y^*_{12}), (X^*_{22}, Y^*_{22}), \ldots, (X^*_{n2}, Y^*_{n2})\right\} \\
\vdots &= \vdots \\
B^*_{B} &= \left\{(X^*_{1B}, Y^*_{1B}), (X^*_{2B}, Y^*_{2B}), \ldots, (X^*_{nB}, Y^*_{nB})\right\} \\
\end{aligned}
\end{equation}


**Step 2**: Let $S^*_{x_b}$ and $S^*_{y_b}$ be the standard deviations of the variables, $X^*_b$ and $Y^*_b$, respectively, where $b = \left\{1, 2, \ldots, B\right\}$. Calculate the pearson product coefficient of correlation, $r^*_b$


\begin{equation}
\begin{aligned}
r_b^* &= \frac{\frac{1}{n-1} \sum \left(X^*_{ib} - {\bar{X}}_b^*\right) \left(Y_{ib}^* - {\bar{Y}}^*_b\right) }{ S^*_{x_b} S^*_{y_b}}
\end{aligned}
\end{equation}


to yield


\begin{equation}
\begin{aligned}
r &= \left\{r^*_1, r^*_2, \ldots, r^*_B\right\} 
\end{aligned}
\end{equation}


**Step 3**: Calculate $\widehat{se}(r)$ using


\begin{equation}
\begin{aligned}
\widehat{se}(r) &= \sqrt{ \frac{\sum_{b=1}^B \left( r_i^* - {\bar{r}}^* \right)^2}{B-1} }
\end{aligned}
\end{equation}


## Algorithm implementation

```{r results = "hide"}
source("../R/s01_i01_load_data.R")
```

```{r se}
file <- "../R/s02_i01_bs_sampling.R"
source(file)
```

```{r algo}
cat(paste(readLines(file), collapse="\n"),collapse = "\n")
```

i.  bootstrap estimate of the standard error of $r$: `r se_r_boot`
ii. 95% confidence interval for $\rho$ (the true population correlation): (`r ci_r_boot`)
iii. a histogram showing the bootstrap distribution of the correlation $r$

```{r}
hist(r)
```
