```{r}
R01_01 <- "../R/01_01_load_data.R"
source(R01_01)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_01, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_01), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Bootstrap-t method: median

Compute a 95% CI for $\theta$ using the bootstrap-$t$ method. Use $B_1 = 1000$ first-level bootstrap samples and $B_2 = 50$ second level bootstrap samples (to estimate the standard error).


```{r}
R01_02 <- "../R/01_02_bst_median.R"
source(R01_02)
```


We are 95% confident that the true value of the median is between `r lower` and `r upper`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_02, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_02), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Bootstrap percentile CI: median

Compute a 95% CI for $\theta$ using the bootstrap percentile CI with B = 1000 bootstrap samples.

```{r}
R01_03 <- "../R/01_03_percentile_median.R"
source(R01_03)
```


We are 95% confident that the true value of the median is between `r quantile(res_median,.025)` and `r quantile(res_median,.975)`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_03, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_03), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Reverse bootstrap percentile CI (aka basic bootstrap): mean

Compute a 95% confidence interval for the mean time between failures $\theta$ using the basic bootstrap method with B = 1000 bootstrap samples.

```{r}
R01_04 <- "../R/01_04_reversepercentile_mean.R"
source(R01_04)
```


We are 95% confident that the true value of the mean is between `r 2*sample_mean - quantile(res_mean,.975)` and `r 2*sample_mean - quantile(res_mean,.025)`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_04, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_04), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Density estimate

Plot a density estimate of the data. In R, you can do this through the density function. Compare
the results in parts 1.1, 1.2, and 1.3.  

For the selected seed, the percentile CI captures the bulk of the median values from the bootstrap samples. The bootstrap-t (median) CI appears to cover areas of large concentration from both the percentile (median) and reverse percentile CI (mean). I also tried a number of seeds and noticed across all cases that: 1) The lower CI endpoint of the basic and the percentile intervals are close to each other. 2) The upper CI endpoint of basic and bootstrap-t intervals are close to each other but the lower CI endpoint of the latter is always less than that of the former. 3) Bootstrap-t captures the bulk of the statistics (mean and median). 4) The upper limit of the percentile interval could be erratic due to the presence of outliers.

```{r}
R01_05 <- "../R/01_05_density_plot.R"
source(R01_05)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_05, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_05), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```















