```{r}
R01_01 <- "../R/01_01_load_data.R"
source(R01_01)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_01, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_01), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Bootstrap-t method: median

Compute a 95% CI for $\theta$ using the bootstrap-$t$ method. Use $B_1 = 1000$ first-level bootstrap samples and $B_2 = 50$ second level bootstrap samples (to estimate the standard error).


```{r}
R01_02 <- "../R/01_02_bst_median.R"
source(R01_02)
```


We are 95% confident that the true value of the median is between `r lower` and `r upper`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_02, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_02), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```



## Bootstrap percentile CI: median

Compute a 95% CI for $\theta$ using the bootstrap percentile CI with B = 1000 bootstrap samples.

```{r}
R01_03 <- "../R/01_03_percentile_median.R"
source(R01_03)
```


We are 95% confident that the true value of the median is between `r quantile(res_median,.025)` and `r quantile(res_median,.975)`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_03, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_03), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Bootstrap percentile CI: mean

Compute a 95% confidence interval for the mean time between failures $\theta$ using the basic bootstrap method with B = 1000 bootstrap samples.

```{r}
R01_04 <- "../R/01_04_percentile_mean.R"
source(R01_04)
```


We are 95% confident that the true value of the mean is between `r quantile(res_mean,.025)` and `r quantile(res_mean,.975)`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_04, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_04), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Density estimate

Plot a density estimate of the data. In R, you can do this through the density function. Compare
the results in parts 1.1, 1.2, and 1.3.  

First, the percentile CI's covers areas in which the the bulk of statistics are seen. Because the data is skewed to the right and median is robust to outliers, percentile CI based on it can be seen to the left of the percentile CI based on the mean. Next, the mean is pulled to the right due to some extremely high values (104, 146). Finally, it could be observed that the percentile CI's are narrower than that of the bootstrap-t, and that the latter appears to cover areas of large concentration from both the mean and the median percentile CI's.

```{r}
R01_05 <- "../R/01_05_density_plot.R"
source(R01_05)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_05, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R01_05), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```















