```{r results = "hide"}
R0201 <- "../R/02_01_load_data.R"
source(R0201)
```

## Bootstrap percentile CI for $\theta_1$

Compute a bootstrap percentile confidence interval for $\theta_1$.

```{r}
R02_02 <- "../R/02_02_percentile_logmean.R"
source(R02_02)
```

We are 95% confident that the true value of $\theta_1$ is between `r theta_1_pci[1]` and `r theta_1_pci[2]`.

```{r}
cat(paste0("  CODE FILENAME: ", R02_02, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R02_02), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```



## $BC_a$ CI for $\theta_1$

Compute a $BC_a$ confidence interval for $\theta_1$. Interpret the CI.

```{r}
R02_03 <- "../R/02_03_bca_logmean.R"
source(R02_03)
```

We are 95% confident that the true value of $\theta_1$ is between `r theta_1_BCa[1]` and `r theta_1_BCa[2]`.

```{r}
cat(paste0("  CODE FILENAME: ", R02_03, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R02_03), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## $BC_a$ CI for $\theta_2$

Compute a $BC_a$ confidence interval for $\theta_2$.

```{r}

```

```{r}
R02_05 <- "../R/02_05_bca_meanlog.R"
source(R02_05)
```

We are 95% confident that the true value of $\theta_2$ is between `r theta_2_BCa[1]` and `r theta_2_BCa[2]`.

```{r}
cat(paste0("  CODE FILENAME: ", R02_05, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R02_05), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```



## Bootstrap percentile vs $BC_a$ CI for $\theta_1$

Compare your CIs in 2.1 and 2.2. How different are the two CIs?

The endpoints of the $BC_a$ CI are found to the right of the bootstrap percentile CI but they are not hugely different. It is possible that the plug-in estimator is a little biased, resulting to that characteristic. Percentile CI works best if $\hat \theta_1$ is an unbiased estimator of $\theta_1$. In our case, we only have bootstrap resamples that are centered around the mean, `r mean(theta_1_hat_star)` which is also a little different from the original sample's plug-in estimate `r plug_in_estimate_theta_1`. $BC_a$ corrects for the bias.

```{r}
R02_06 <- "../R/02_06_compare_ci.R"
source(R02_06)
```


```{r}
cat(paste0("  CODE FILENAME: ", R02_06, paste(rep("*",5), collapse = "")))
cat(paste(readLines(R02_06), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```