---
title: "May5"
author: "Shaine"
date: "2023-05-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recall

Bootstrap percentile CI where, $\theta^*_1, \dots, \theta^*_B \rightarrow (1-\alpha)100\%$ bootstrap percentile CI for $\theta$ is $\left( \hat \theta^*_\frac{\theta}{2}, \hat \theta^*_{1 -\frac{\theta}{2}} \right)$ (quantiles of bs statistics)

This works best if $\hat \theta$ is an unbiased estimator of $\theta$. It is problematic when the bias is large. We want $\hat \theta$ to be an unbiased estimator of $\theta$ since the bs distribution is centered around $\hat \theta$ (statistic computed from our sample). The reason it is centered around $\hat \theta$ is that these are just resamples (taken from the sample).

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
