```{r micro,results = "hide"}
R01_01 <- "../R/01_01_bst.R"
source(R01_01)
```

```{r results = 'asis'}
stargazer::stargazer(t(data.frame(data)), 
          header=FALSE, type='latex',
          title="Survival Times (Days) for Patients with Breast Cancer",
          label = "tab:original",
          rownames = FALSE)
```

## Bootstrap-t

Use the bootstrap-$t$ method to construct $95 \%$ CI for the mean survival time. *When doing the bootstrap, work with the data on the natural log scale and then exponentiate the resulting interval boundaries.* Interpret the resulting interval.

We are 95% confident that the true value of mean survival time is between `r exp(lower)` and `r exp(upper)`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_01, paste(rep("-", 43), collapse = "")))
cat(paste(readLines(R01_01), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```


## BCa

Use the $BCa$ method to construct a $95 \%$ CI for the mean survival time.


```{r}
R01_02 <- "../R/01_02_bca.R"
source(R01_02)
```

We are 95% confident that the true value of mean survival time is between `r exp(theta_BCa)[1]` and `r exp(theta_BCa)[2]`.

```{r}
cat(paste0("  CODE FILENAME: ", R01_02, paste(rep("-", 45), collapse = "")))
cat(paste(readLines(R01_02), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```

## Comparison

Compare your intervals.


Both intervals (shown below - NOT yet exponentiated) for $BSt$ and $BCa$ contain the sample mean survival time, `r sample_mean`. It is a bit wider for $BSt$ (upper - lower: `r upper-lower`) than $BCa$ (upper - lower: `r theta_BCa[2]-theta_BCa[1]`). Upon exponentiation, the difference between the two CI's became larger. $BSt$ (upper - lower: `r exp(upper)-exp(lower)`) is hugely wider than that of $BCa$ (upper - lower: `r exp(theta_BCa[2])-exp(theta_BCa[1])`). 

Both of these CI's are second order accurate but $BCa$ is range preserving and transformation respecting, unlike $BSt$

```{r}
R01_03 <- "../R/01_03_compare.R"
source(R01_03)
```

```{r}
cat(paste0("  CODE FILENAME: ", R01_03, paste(rep("-", 45), collapse = "")))
cat(paste(readLines(R01_03), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```


