
```{r}
R02_01 <- "../R/02_01_bstest.R"
source(R02_01)
```

```{r results = 'asis'}
X <- data.frame("Subject" = 1:8, 
"Baseline" = caffeine_data$baseline, 
"Caffeine" = caffeine_data$caffeine)

row.names(X) <- NULL

stargazer::stargazer(X, type='latex',
                     header=FALSE,
          title="MBF (ml/min/g) for eight subjects",
          label = "tab:original",
          summary = FALSE,
          rownames = FALSE)
```



## Bootstrap test

For this kind of data, explain how you would perform a bootstrap hypothesis test to check if there is a difference in mean MBF.

Some notes:

- Since this is a test not on the distribution, but on the value of the parameter, the bootstrap test will be used.  
- Since the observations are paired or matched, samples will be taken pairwise.  
- Test the hypothesis that the population's average difference in MBF levels during exercise after taking nothing - baseline($X$), and taking 200 mg caffeine($Y$) is equal to zero.

\begin{equation}
\begin{aligned}
H_0: D = \sum_{i=1}^N\left(x_i-y_i\right) = 0 \\
H_1: D = \sum_{i=1}^N\left(x_i-y_i\right) \neq 0
\end{aligned}
\end{equation}

To administer the bootstrap hypothesis test,

1. Transform the data so that it follows the null hypothesis. 

   Given $X$ and $Y$, the differences of pairs are given by
   
   \begin{equation}
   \begin{aligned}
   d_i = x_i - y_i; \quad i = 1,\dots, n
   \end{aligned}
   \end{equation}
   
   The following transformation is applied so that the null hypothesis applies to the sample.
   
   \begin{equation}
   \begin{aligned}  
   \tilde {d_i} = d_i - \bar d;  \quad \bar d = \frac{\sum_{i=1}^nd_i}{n}
   \end{aligned}
   \end{equation}
   
   Below is the proof that this transformation is guaranteed to result in 0.
   
   \begin{equation}
   \begin{aligned}
   \frac{1}{n}\sum_{i=1}^n \tilde {d_i} 
   = \frac{1}{n}\sum_{i=1}^n \left(d_i - \bar d \right) 
   = \frac{1}{n}\sum_{i=1}^n d_i - \frac{1}{n}\sum_{i=1}^n \bar d
   = \frac{1}{n}\sum_{i=1}^n d_i - \frac{1}{n}n\bar d 
   = \bar d - \bar d 
   = 0
   \end{aligned}
   \end{equation}

2. Calculate the test statistic for the observed data: t-statistic for paired or matched samples

   \begin{equation}
   \begin{aligned}
   t_{obs} = \frac{\bar {\tilde d}}{\frac{s}{\sqrt n}}; \quad s = \sqrt{\frac{\sum_{i=1}^n \left(\tilde d - \bar{\tilde d}\right)^2}{n-1}}
   \end{aligned}
   \end{equation}


3. Take B bootstrap samples by sampling pairs (or matched differences), with replacement, of size n from the transformed data. Let $b = 1, \dots, B$

   \begin{equation}
   \begin{aligned}
   B_1^* = \{\tilde d_{11}^*, \dots, \tilde d_{1n}^*\} \overset {SRSWR}\sim \{\tilde d_1, \dots, \tilde d_n\} \\
   B_2^* = \{\tilde d_{21}^*, \dots, \tilde d_{2n}^*\} \overset {SRSWR}\sim \{\tilde d_1, \dots, \tilde d_n\} \\
   \vdots \\
   B_B^* = \{\tilde d_{B1}^*, \dots, \tilde d_{Bn}^*\} \overset {SRSWR}\sim \{\tilde d_1, \dots, \tilde d_n\} \\
   \end{aligned}
   \end{equation}

4. Compute the test statistic for each bootstrap sample. This gives $B$ values.

   \begin{equation}
   \begin{aligned}
   t_b^* = \frac{\bar {\tilde d^*}}{\frac{s^*}{\sqrt n}}; \quad s^* = \sqrt{\frac{\sum_{i=1}^n \left(\tilde d^* - \bar{\tilde d^*}\right)^2}{n-1}}
   \end{aligned}
   \end{equation}
   
5. Estimate the p-value. If it is less than the significance level ($\alpha$), $0.05$, reject the null hypothesis.

   \begin{equation}
   \begin{aligned}
   P(|t^*| \geq |t_{obs}|) = \frac{ \# \{ |t^*| \geq |t_{obs}| \}}{B}
   \end{aligned}
   \end{equation}

## Implementation

Perform the bootstrap hypothesis test at $\alpha =0.05$. 

With the estimated p-value of `r sum(indicator)/B > 0.05`, there is not enough evidence to say that there is a difference in the average MBR levels when taking 200 mg caffeine and not, during exercise.

```{r}
cat(paste0("  CODE FILENAME: ", R02_01, paste(rep("-", 43), collapse = "")))
cat(paste(readLines(R02_01), collapse="\n"),collapse = "\n")
cat(paste0("  ", "end",paste(rep("-", 73), collapse = "")))
```
