---
number_sections: true
fig_caption: yes
output:
  bookdown::pdf_document2:
    toc: false
  citation_package: natbib
fontsize: 12pt 
header-includes:
- \usepackage{amsmath}
- \numberwithin{equation}{section}
- \usepackage{setspace}\singlespacing
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{titling}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[RO,RE]{Matala}
- \fancyhead[LO,LE]{Homework 2}
- \fancyfoot[LO,LE]{`r format(Sys.time(), 'Last rendered on %Y %B %d %X')`}
- \fancyfoot[CO,CE]{}
- \fancyfoot[RE,RO]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
geometry: margin=2cm
papersize: a4
bibliography: /home/scientists/sci01/Projects/bootstrap/common/ref.bib
biblio-style: apalike
csl: /home/scientists/sci01/Projects/bootstrap/common/harvard-educational-review.csl
nocite: |
  @tibs
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(
  echo = FALSE, 
  comment = NA, 
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center',
  fig.height = 4.5,
  fig.width = 5.5)
options(scipen = 1, digits = 5)
```

# Rats

As a small example, the survival times of 9 rats were 10, 27, 30, 40, 46, 51, 52, 104, and 146 days. Because of the skewness in the data, consider estimating the population median survival time $\theta$ through the sample median.




```{r, child = "child/item_01.Rmd"}

```

# Spatial test

Consider the spatial test data from Table 14.1 of Efron and Tibshirani (1993) shown below. From the table's description, it is clear that the measurements A and B are paired. Suppose the data consist of a random sample from an unknown joint distribution of A and B. Whenever ratios are scientifically or statistically preferred to differences, we gain stability by considering the logarithm of the ratios. Let $\theta_1 = \log E \left(\frac{A_i}{B_i}\right)$, $\theta_2 = E \left(\log \frac{A_i}{B_i}\right)$ for all $i$. Exclude observation #14 because the logarithm of its ratio is undefined. Use 2000 bootstrap samples.

```{r, child = "child/item_02.Rmd"}

```

# References

<div id="refs"></div>

# Appendix

```{r, child = "child/appendix.Rmd"}

```